Einstein's gravity from a polynomial affine model

### 2016-04-08 AffGR-PRD.tex
Introduction reviewed

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds ---with or without cosmological constant--- as solutions. Moreover, these effective field equations coincide with some of those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model. Finally, we present a naive analysis of the propagators of the gravitational theory, an their relation with the presence of ghost in the model. 
\end{abstract}

@@

However, General Relativity seems to be an effective theory of a, yet to be discover, more fundamental one. A first hint was pointed out by E.~Cartan, after noticing that that a part of the connection was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's framework for General Relativity, in which both fields are independent, and noticing that the theory is not complete because there are terms which are zero in pure General Relativity that do not vanish in Palatini's generalization. In addition, we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions ---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, and Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Standing on the argument of the necessity of a quantum theory of gravity, several generalizations of Einstein's General Relativity have been proposed. One important class of these generalizations considers the connection field to be independent of the metric, known as Palatini type theories. Within this class of theories one finds the generalization due to Cartan, which considers a connection with nonvanishing torsion~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. Interestingly, gravitational theories with generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe} or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}.

%%%%%%%%% THE MODEL %%%%%%%%%
In order to undertake the premises of incompleteness and non-polynomial structure of the theory, in a previous paper~\cite{Skirzewski:2014eta}, we presented a polynomial (purely) affine gravitational model in four dimensions built up entirely on the basis of full diffeomorphism invariance. Worth to mention that the most general action built this way contains only power-counting renormalizable terms (which by no means guarantees the renormalizability of the system). We showed that its non-relativistic limit around a homogeneous and isotropic spacetime yields a Newtonian gravitational potential, despite the existence of torsion. We also highlighted the possibility of using standard methods to quantize the model, due to its polynomial structure (unlike the earliest proposals~\cite{Eddington1923math,schrodinger1950space}), and also the likelihood of avoiding the uniqueness of diffeomorphism-invariant states from Loop Quantum Gravity programme~\cite{Lewandowski:2005jk}, due to the absence of a fundamental metric, or in other words, the absence of flux operators.

Interestingly, by construction, the polynomial affine gravity proposed in Ref.~\cite{Skirzewski:2014eta} has no explicit terms leading to three-point graviton vertices, since all graviton self-interaction is mediated by the torsion. This characteristic might allow to bypass the general postulates supporting the no-go theorems stated in Refs.~\cite{McGady:2013sga,Camanho:2014apa}, where it is proved that generic three-point graviton interactions are highly constrained by causality and analyticity of the $S$-matrix, and the only \emph{acceptable} structure of the three-point graviton vertices is the one coming from General Relativity.
%%Additionally, it is well known that the three-point graviton vertices lead to non-renormalizability of a theory ---in a model independent way---~\cite{McGady:2013sga,Camanho:2014apa}. However, in the considered framework all graviton self-interaction is mediated by the torsion, which might allow to bypass the general postulates supporting the theorem.

The aim of this work is to show that in the vanishing torsion sector, the field equations of the polynomial affine gravity are a generalization of the Einstein's equations. % in vacuum. In addition, we show that ---in the same sector--- under the assumption of the existence of a unique non-degenerated, covariantly constant $\binom{0}{2}$-tensor, $g$ (which could be thought as a metric), the system of a scalar field coupled with the polynomial affine gravity accepts the minimally coupled Einstein--Klein--Gordon as a solution.
The paper is organized as follows. In Sec.~\ref{model}, we present briefly the polynomial affine gravity model, and summarize the results obtained in Ref.~\cite{Skirzewski:2014eta} concerning the Newtonian limit of the model. In Sec.~\ref{rlimit}, we restrict ourselves to the sector of the theory with vanishing torsion, and find the field equations; which surprisingly are a known generalization of the Einstein's field equation. Then, in Sec.~\ref{matter} we show that under certain considerations the coupling of the gravitational model with a scalar field represents a generalization of the minimally coupled system Einstein--Klein--Gordon. Finally, in Sec.~\ref{conclusions} we summarize and discuss the possible implications of the results.

@@

%%%%%%%%% Propagators %%%%%%%%%
\section{Analysis of propagators}

>>>>

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds ---with or without cosmological constant--- as solutions. Moreover, the effective field equations coincide with some of those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model. Finally, we present a naive analysis of the propagators of the gravitational theory, an their relation with the presence of ghost in the model. 
\end{abstract}

@@

However, General Relativity seems to be an effective theory of a, yet to be discover, more fundamental one. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity. In addition, we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions ---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Standing on the argument of the necessity of a quantum theory of gravity, several generalizations of Einstein's General Relativity have been proposed. One important class of these generalizations considers the connection field to be independent of the metric, known as Palatini type theories. Within this class of theories one finds the generalization due to Cartan---which considers a connection with nonvanishing torsion~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}---, but also a broader kind of generalization which do not require the metricity condition, see for example Refs.~\cite{Hehl:1976kj,Hehl:1994ue,Pagani:2015ema}.

Interestingly, gravitational theories with generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe} or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}. The last sentence, allows to argue that a model cannot be discarded  for the presence of ghosts model, without a rigorous analysis of its stability.

%%%%%%%%% THE MODEL %%%%%%%%%
In order to undertake the premises of incompleteness and non-polynomial structure of the theory, in a previous paper~\cite{Skirzewski:2014eta}, we presented a polynomial (purely) affine gravitational model in four dimensions built up entirely on the basis of full diffeomorphism invariance. Worth to mention that the most general action built this way contains only power-counting renormalizable terms (which by no means guarantees the renormalizability of the system). We showed that its non-relativistic limit, around a homogeneous and isotropic spacetime, yields a geodesic deviation corresponding to the generated by a Newtonian gravitational potential, despite the existence of torsion. We also highlighted the possibility of using standard methods to quantize the model, due to its polynomial structure (unlike the earliest proposals~\cite{Eddington1923math,schrodinger1950space}), and also the likelihood of avoiding the uniqueness of diffeomorphism-invariant states from Loop Quantum Gravity programme~\cite{Lewandowski:2005jk}, due to the absence of a fundamental metric, or in other words, the absence of flux operators.

Interestingly, by construction, the polynomial affine gravity proposed in Ref.~\cite{Skirzewski:2014eta} has no explicit terms leading to three-point graviton vertices, since all graviton self-interaction is mediated by the torsion. This characteristic might allow to bypass the general postulates supporting the no-go theorems stated in Refs.~\cite{McGady:2013sga,Camanho:2014apa}, where it was proved that generic three-point graviton interactions are highly constrained by causality and analyticity of the $S$-matrix, and the only \emph{acceptable} structure of the three-point graviton vertices is the one coming from General Relativity.
%%Additionally, it is well known that the three-point graviton vertices lead to non-renormalizability of a theory ---in a model independent way---~\cite{McGady:2013sga,Camanho:2014apa}. However, in the considered framework all graviton self-interaction is mediated by the torsion, which might allow to bypass the general postulates supporting the theorem.

The aim of this work is to show that in the vanishing torsion sector, the field equations of the polynomial affine gravity are a generalization of the Einstein's equations. % in vacuum. In addition, we show that ---in the same sector--- under the assumption of the existence of a unique non-degenerated, covariantly constant $\binom{0}{2}$-tensor, $g$ (which could be thought as a metric), the system of a scalar field coupled with the polynomial affine gravity accepts the minimally coupled Einstein--Klein--Gordon as a solution.
The paper is organized as follows. In Sec.~\ref{model}, we present briefly the polynomial affine gravity model, and summarize the results obtained in Ref.~\cite{Skirzewski:2014eta} concerning the Newtonian limit of the model. In Sec.~\ref{rlimit}, we restrict ourselves to the sector of the theory with vanishing torsion, and find the field equations; which surprisingly are a known generalization of the Einstein's field equation. Then, in Sec.~\ref{matter} we show that under certain considerations the coupling of the gravitational model with a scalar field represents a generalization of the minimally coupled system Einstein--Klein--Gordon. In Sec.~\ref{propag}, we present a brief analysis of the propagators of the model, which represent a first step toward a thorough study of the degrees of freedom of these theory. And in Sec.~\ref{conclusions} we summarize and discuss the possible implications of the results. In order to make the paper self-contained, we have included three appendices: In Appendix~\ref{DA}, we describe a dimensional analysis strategy that allowed us to build up the most general action consistent with diffeomorphisms. In Appendix~\ref{Known}, we show a collection of known solutions to the effective field equations of the model. Finally, in Appendix~\ref{Schw}, we demonstrate---restricting ourselves to the vanishing torsion sector---the equivalent of the Birkhoff's theorem of the theory.

@@

%%%%%%%%% Propagators %%%%%%%%%
\section{\label{propag}Analysis of propagators}

### 2016-04-08 MS.tex
Introduction reviewed

\section{Known metric solutions to the parallel Ricci equations}

@@

\section{Explicit example: Schwarzschild ansatz, and (incomplete) Birkhoff theorem\label{Ex:sch}}

@@

@@

As argue  in Refs.~\cite{bergmann1965spherically,hawking1973large},  the solution depends on the nature of the surfaces described by $\rho = \text{const.}$, which in our parametrization translate to the nature of $N_a$. We first consider the condition $g^{ab} N_a N_b > 0 $, which allows to set $\rho$ as a good radial coordinate. Therefore, the general metric ansatz $g_{ab}$ is diagonal, and the radial direction would be $\rho$, i.e.,
\begin{equation}
  g_{ab} = - \alpha^2(t,\rho) \de{t}_a \otimes \de{t}_b + \beta^2(t,\rho) \de{\rho}_a \otimes \de{\rho}_b.
\end{equation}
With this ansatz, the nontrivial components of Eq.~\eqref{OurEq} are
\begin{align}
  \beta^2 \pa{t} \mathcal{R} + 8 \gamma^\rho{}_{t \rho} &= 0,
  \label{eq1}
  \\
  \pa{\rho} \mathcal{R} - 4 \mathcal{R} + 16 \alpha^{-2} \gamma^\rho{}_{tt} &= 0,
  \label{eq2}
  \\
  \pa{t} \big( \Delta_x e^{2\rho} \big) + 4 \beta^{-2} e^{2\rho} \gamma^\rho{}_{t \rho} &= 0,
  \label{eq3}
\end{align}
and
\begin{dmath}
  \pa{\rho} \big( \Delta_x e^{2\rho} \big) + 8 - 4 \Delta_x e^{2\rho} - 16 \beta^{-2} e^{2\rho} - 4 \beta^{-2} e^{2\rho} \gamma^{\rho}{}_{\rho\rho}  + \mathcal{R} e^{2\rho} = 0.
  \label{eq4}
\end{dmath}
Equation~\eqref{eq1} can be rewritten as
\begin{equation}
  \pa{t} \big( \mathcal{R} - 8 \beta^{-2} \big) = 0,
\end{equation}
or 
\begin{equation}
  \mathcal{R} - 8 \beta^{-2} = G_1(\rho).
\end{equation}
Since
\begin{equation}
  \Delta_x e^{2\rho} = e^{2\rho} \bigg[ 4 \beta^{-2} + \frac{2}{\alpha \beta} \pa{\rho} \bigg( \frac{\alpha}{\beta} \bigg) \bigg] \equiv  e^{2\rho} \Big[ 4 \beta^{-2} + F \Big],
\end{equation}
it follows from Eq.~\eqref{eq3} that,
\begin{equation}
  mmmmmmmmm
\end{equation}

>>>>

\section{\label{Known}Known metric solutions to the parallel Ricci equations}

@@

\section{\label{Schw}Explicit example: Schwarzschild ansatz, and Birkhoff theorem\label{Ex:sch}}

@@

    \label{hugeeq}

@@

As argue  in Refs.~\cite{bergmann1965spherically,hawking1973large},  the solution depends on the nature of the surfaces described by $\rho = \text{const.}$, which in our parametrization translate to the nature of $N_a$. We first consider the condition $g^{ab} N_a N_b > 0 $, which allows to set $\rho$ as a good radial coordinate. Therefore, the general metric ansatz $g_{ab}$ is diagonal, and the radial direction would be $\rho$, i.e.,
\begin{equation}
  g_{ab} = - \alpha^2(t,\rho) \de{t}_a \otimes \de{t}_b + \beta^2(t,\rho) \de{\rho}_a \otimes \de{\rho}_b.
  \label{metr2d}
\end{equation}
With this ansatz, the nontrivial components of Eq.~\eqref{OurEq} are
\begin{align}
  \beta^2 \pa{t} \mathcal{R} + 8 \gamma^\rho{}_{t \rho} &= 0,
  \label{eq1}
  \\
  \pa{\rho} \mathcal{R} - 4 \mathcal{R} + 16 \alpha^{-2} \gamma^\rho{}_{tt} &= 0,
  \label{eq2}
  \\
  \pa{t} \big( \Delta_x e^{2\rho} \big) + 4 \beta^{-2} e^{2\rho} \gamma^\rho{}_{t \rho} &= 0,
  \label{eq3}
\end{align}
and
\begin{dmath}
  \pa{\rho} \big( \Delta_x e^{2\rho} \big) + 8 - 4 \Delta_x e^{2\rho} - 16 \beta^{-2} e^{2\rho} - 4 \beta^{-2} e^{2\rho} \gamma^{\rho}{}_{\rho\rho}  + \mathcal{R} e^{2\rho} = 0.
  \label{eq4}
\end{dmath}
From equation~\eqref{eq1} it follows that,
%% can be rewritten as
%% \begin{equation}
%%   \pa{t} \big( \mathcal{R} - 8 \beta^{-2} \big) = 0,
%%   \label{eq1a}
%% \end{equation}
%% or 
\begin{equation}
  \mathcal{R} - 8 \beta^{-2} = G_1(\rho).
  \label{eq1b}
\end{equation}
In addition, since
\begin{equation}
  \Delta_x e^{2\rho}
  = e^{2\rho} \bigg[ 4 \beta^{-2} + \frac{2}{\alpha \beta} \pa{\rho} \bigg( \frac{\alpha}{\beta} \bigg) \bigg]
  \equiv  e^{2\rho} \Big[ 4 \beta^{-2} + \mathcal{F} \Big],
\end{equation}
it follows from Eqs.~\eqref{eq3} that,
%% \begin{equation}
%%   \pa{t} \big( F + 2 \beta^{-2} \big) = 0,
%%   \label{eq3a}
%% \end{equation}
%% or 
\begin{equation}
  \mathcal{F} + 2 \beta^{-2} = G_2(\rho).
  \label{eq3b}
\end{equation}

Equations~\eqref{eq1b} and~\eqref{eq3b} implies that all temporal dependence on time within $\mathcal{R}$, $\mathcal{F}$ and $\beta^{-2}$, enters through an additive common function, say $B(t,\rho)$. However, the consistency of these equations with the value of $\mathcal{R}$ calculated from the metric ansatz, in Eq.~\eqref{metr2d}, sets this additive function to zero, $B(t,\rho)$. Therefore, neither $\mathcal{R}$, $\mathcal{F}$ nor $\beta^{-2}$ depend on $t$. By extension---from the definition of $\mathcal{F}$---, $\alpha$ could depend on time through a multiplicative function. Nonetheless, this multiplicative function can be eliminated of the problem by a redefinition of the \emph{time} coordinate.

At this stage, we have proved that \emph{any spherically symmetric (metric) solution of the Eqs.~\eqref{SimpleEOM} is necessarily static}, which is one of the main statements of Birkhoff's theorem. This result, simplifies the Eq.~\eqref{hugeeq} and makes clearer the necessity of setting $A(r) = B(r)$ to solve the equations of motion. Finally, according to Eq.~\eqref{simplef} the general (metric) solution is the (Anti-)de-Sitter--Schwarzschild metric.
